---
title: "Proj2_154"
author: "Sizhuo (Cindy) Liu"
date: "4/20/2019"
output: pdf_document
---

1. Data Collection and Exploration

The purpose of the study is to to propose 

```{r}
# Load Image Data & Add Column Names
image1 <- read.table("image_data/image1.txt")
colnames(image1) <- c('y_coordinate', 'x_coordinate', 'expert_label', 'NDAI', 'SD', 'CORR', 'RA_DF', 'RA_CF', 'RA_BF', 'RA_AF', 'RA_AN')
image1$expert_label <- as.factor(image1$expert_label)
image2 <- read.table("image_data/image2.txt")
colnames(image2) <- c('y_coordinate', 'x_coordinate', 'expert_label', 'NDAI', 'SD', 'CORR', 'RA_DF', 'RA_CF', 'RA_BF', 'RA_AF', 'RA_AN')
image2$expert_label <- as.factor(image2$expert_label)
image3 <- read.table("image_data/image3.txt")
colnames(image3) <- c('y_coordinate', 'x_coordinate', 'expert_label', 'NDAI', 'SD', 'CORR', 'RA_DF', 'RA_CF', 'RA_BF', 'RA_AF', 'RA_AN')
image3$expert_label <- as.factor(image3$expert_label)

```

```{r}
# Plot Well-Labeled Map
library(ggplot2)
ggplot(data = image1) + geom_point(aes(x = x_coordinate, y = y_coordinate, color = as.factor(expert_label))) + scale_y_reverse() + scale_color_manual("Expert Labels", labels = c("Not Cloud", "No Label", "Cloud"), values = c("grey", "black", "white"))
ggplot(data = image2) + geom_point(aes(x = x_coordinate, y = y_coordinate, color = as.factor(expert_label))) + scale_y_reverse() + scale_color_manual("Expert Labels", labels = c("Not Cloud", "No Label", "Cloud"), values = c("grey", "black", "white"))
ggplot(data = image3) + geom_point(aes(x = x_coordinate, y = y_coordinate, color = as.factor(expert_label))) + scale_y_reverse() + scale_color_manual("Expert Labels", labels = c("Not Cloud", "No Label", "Cloud"), values = c("grey", "black", "white")) + scale_color_manual("Expert Labels", labels = c("Not Cloud", "No Label", "Cloud"), values = c("grey", "black", "white"))
```

```{r}
# Observe pair-wise relationships between variables ??? Plots different from in the paper
library(scales)
pairs(image1_nozero[,c(4, 5, 6)])
pairs(image2_nozero[,c(4, 5, 6)])


image1_nozero <- image1[-which(image1$expert_label == 0),]
rownames(image1_nozero) <- 1:nrow(image1_nozero)
ggplot(data = image1_nozero) + geom_point(aes(x = CORR, y = rescale(image1_nozero$NDAI, to = c(-0.5, 1)), color = as.factor(expert_label))) + scale_color_manual("Expert Labels", labels = c("Not Cloud", "Cloud"), values = c("blue", "red")) + coord_flip()

image2_nozero <- image2[-which(image2$expert_label == 0),]
rownames(image2_nozero) <- 1:nrow(image2_nozero)
ggplot(data = image2_nozero) + geom_point(aes(x = CORR, y = NDAI, color = as.factor(expert_label))) + scale_color_manual("Expert Labels", labels = c("Not Cloud", "Cloud"), values = c("blue", "red"))

image3_nozero <- image3[-which(image3$expert_label == 0),]
rownames(image3_nozero) <- 1:nrow(image3_nozero)
ggplot(data = image2_nozero) + geom_histogram(aes(x = rescale(image2_nozero$NDAI, to = c(-0.5, 1))))

```

```{r}
# Pairwise Relationships with Expert Label 
ggplot(data = image1_nozero) + geom_boxplot(aes(y = CORR, color = expert_label))
boxplot(image1_nozero$CORR)

ggplot(data = image1_nozero) + geom_histogram(aes(x = CORR, color = expert_label, fill = expert_label, alpha = 0.5))

ggplot(data = image1_nozero) + geom_histogram(aes(x = SD, color = expert_label, fill = expert_label, alpha = 0.5)) + scale_color_manual("Expert Labels", labels = c("Not Cloud", "Cloud"), values = c("blue", "red"))

ggplot(data = image1_nozero) + geom_histogram(aes(x = NDAI, color = expert_label, fill = expert_label, alpha = 0.5)) + scale_color_manual("Expert Labels", labels = c("Not Cloud", "Cloud"), values = c("blue", "red"))
```

2. Preparation

```{r}
# a) Data Split

# Method 1: Split the data from each image individually and then combine 
# Andrew Ng's Recommendation: Training: 60%, Validation: 20%, Testing: 20%
# Image1 Split

split_set <- function(data, num) {
    data$image <- rep(num, nrow(data))
    all_indices <- 1:nrow(data)
    test_index <- sample(all_indices, nrow(data) * 0.2)
    remaining_index <- all_indices[-test_index]
    training_index <- sample(remaining_index, length(remaining_index) * 0.8)
    validation_index <- remaining_index[-training_index]
    train <- data[training_index,]
    validation <- data[validation_index,]
    test <- data[test_index,]
    return(list(train = train,validation = validation,test = test))
}

data1 <- split_set(image1_nozero, 1)
data2 <- split_set(image2_nozero, 2)
data3 <- split_set(image3_nozero, 3)
train <- rbind(data1$train, 
               data2$train, data3$train)

validation <- rbind(data1$validation,
                    data2$validation,
                    data3$validation)

test <- rbind(data1$test,
                    data2$test,
                    data3$test)

```



```{r}
# Method 2: Combine the data and then split randomly 
images <- rbind(image1_nozero, image2_nozero, image3_nozero)
index_all <- 1:nrow(images)
test_indices <- sample(index_all, nrow(images) * 0.2)
remaining_indices <- index_all[-test_indices]
training_indices <- sample(remaining_indices, length(remaining_indices) * 0.8)
validation_indices <- remaining_indices[-training_indices]

train_M2 <- images[training_indices,]
validation_M2 <- images[validation_indices,]
test_M2 <- images[test_indices,]
table(train_M2$image) / sum(table(train_M2$image))

```

b)
The trivial classifier will have a high accuracy when the original image has mostly clear areas (without clouds).
- Method 1 is preferred over Method 2.
- The way method 1 is constructed ensures that in training, validation and test sets, there are equal amounts of data from each image. 
- The table at the end of Method 2, however, suggests that the random selection tends to contain more data from one image than another.

```{r}
baseline_preds <- -1
# validation accuracy
mean(baseline_preds == validation$expert_label)
# test accuracy
mean(baseline_preds == test$expert_label)
image1_nozero
```

```{r}
# c) Ask??? Can we use NDAI, CORR, SD?

```


```{r}
# d)
CVgeneric <- function(classifier, features, labels, K, loss) {
  library(caret)
  folds <- createFolds(labels, k = K)
  error <- rep(0, K)
  for (i in 1: K) {
    test <- features[folds[[i]],]
    train <- features[-folds[[i]],]
    label.train <- labels[folds[[i]]]
    label.test <- labels[-folds[[i]]]
    train$label <- label.train
    # full.test <- cbind(test, label.test)
    # features_added <- paste(colnames(features), collapse = " + ")
    # formula <- paste('labels', ' ~ ', features_added)
    fit <- glm(label ~ SD + CORR + NDAI, family = binomial(link = 'logit'),data=train)
    y_hat <- predict(fit, test, type = 'response')
    error[i] <- loss(label.test, y_hat)
  }
  return (error)
}




```

3. Modeling
```{r}
# Logistic Regression Model 

# Split 1
train_validation <- rbind(train, validation)
train_validation
logistic_mod1 <- glm(expert_label ~ SD + CORR + NDAI, 
                    family = binomial(link = 'logit'), data = train_validation)

logistic_loss <- function(labels, y_hat) {
  return (mean(log(1 + exp(-as.integer(labels)*y_hat))))
}

CVgeneric(logistic_mod1, train_validation[,-3], train_validation$expert_label, K = 5, logistic_loss)

```

```{r}
# test
classifier <- logistic_mod1
features <- train_validation[,-3]
labels <- train_validation$expert_label
K <- 5
loss <- logistic_loss

  library(caret)
  folds <- createFolds(labels, k = K)
  error <- rep(0, K)
  for (i in 1: K) {
    test <- features[folds[[i]],]
    train <- features[-folds[[i]],]
    label.train <- labels[folds[[i]]]
    label.test <- labels[-folds[[i]]]
    train$label <- label.train
    # full.test <- cbind(test, label.test)
    # features_added <- paste(colnames(features), collapse = " + ")
    # formula <- paste('labels', ' ~ ', features_added)
    fit <- glm(label ~ SD + CORR + NDAI, family = binomial(link = 'logit'),data=train)
    y_hat <- predict(fit, test, type = 'response')
    error[i] <- loss(label.test, y_hat)
  }
```



















